IML Term project

### 2022-12-07 main_markdown.rmd
Reordered things.

# Set-up and data preprosessing

@@

# Find cross-validation predictions
cv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
loocv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:nrow(data)) {
        ## go through all stations, train on other stations, and make a prediction
        mod <- train(data[-i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[i] <- pred(mod, data[i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
```

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

@@

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

#Thus, we come up to the following collection of variables
coefficients(testmod)
colnames(npf)
npfsp<-npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact<-npfsp
npfsp.fact[,19]<-as.factor(npfsp.fact[,19])
```

### PCA
#PCA for all three versions of data 
```{r PCA}
pcaOrig <- prcomp(npf[,1:100], center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[,1:100])%*%pcaOrig$rotation[,1:12]
pca.168 <- prcomp(npf.168[,1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp <- prcomp(npfsp[,1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]
```

@@

>>>>

@@

@@

# Find cross-validation predictions, return probabilities and not the classification
cv.prob <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}

loocv.prob <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:nrow(data)) {
        ## go through all stations, train on other stations, and make a prediction
        mod <- train(data[-i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[i] <- pred(mod, data[i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
```

@@

testmod <- glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact <- vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
```

Runnig vif() over several variables in the 16.8m-restricted dataset, we come up to the following collection of coefficients 
```{r vif coefficients}
coefficients(testmod)
```

Hence we keep the following parameters
```{r vif parameters}
npfsp<-npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact<-npfsp
npfsp.fact[,19]<-as.factor(npfsp.fact[,19])
```

### PCA
#### PCA for all three versions of data 
```{r PCA}
pcaOrig <- prcomp(npf[,1:100], center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1 <- as.matrix(npf[,1:100])%*%pcaOrig$rotation[,1:12]
pca.168 <- prcomp(npf.168[,1:36], center = T, scale. = T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2 <- as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp <- prcomp(npfsp[,1:18], center = T, scale. = T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3 <- as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]
```

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv.prob(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran

nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```
